---
title: "Term paper - Empirical industrial economics"
author: "Coraline Best"
date: "2023-02-21"
output: pdf_document
---

# Part 1: Cleaning of the datasets

***1) Cleaning of the economic characteristics dataset***

```{r setup, include=FALSE}
#Loading packages
library(pacman)
p_load(tidyverse, data.table, broom, parallel, here, plotly, ggplot2, stargazer, magrittr,skimr,janitor,  tidymodels, ADAPTS, caret, yardstick, rlang, parsnip, sandwich, lmtest, haven, tinytex, rdrobust,dplyr, plotrix, plyr,readxl, usmap, stringr, finalfit, scales,tidyr, gridExtra, patchwork, EventStudy, fixest,kableExtra,wesanderson, gtsummary)

#Setting directory
setwd("/Users/coraline/Desktop/Econometrics project")

#Importation of the different datasets of Economic characteristics from 2010 to 2017
files_e<- c("2010.SelectedEconomicCharacteristics.xlsx","2011.SelectedEconomicCharacteristics.xlsx", "2012.SelectedEconomicCharacteristics.xlsx","2013.SelectedEconomicCharacteristics.xlsx", "2014.SelectedEconomicCharacteristics.xlsx", "2015.SelectedEconomicCharacteristics.xlsx","2016.SelectedEconomicCharacteristics.xlsx", "2017.SelectedEconomicCharacteristics.xlsx" )

#Assign names to the datasets:
names<-c("e2010","e2011","e2012","e2013","e2014","e2015","e2016","e2017")

#open all the datasets using a loop
for (i in seq_along(files_e)) {
  assign(names[i], read_excel(path = files_e[i], sheet=2 ))
}


#merging all the dataset
economic_cha<- cbind.data.frame(e2010,e2011,e2012,e2013,e2014,e2015,e2016,e2017)

#keeping the variable we need
economic_cha  <- economic_cha[c(6,7, 58:68),]

#transposing data
economic_cha <- data.frame(t(economic_cha), stringsAsFactors=FALSE)
names(economic_cha) <- economic_cha[1,]
economic_cha<- economic_cha[-1,]
economic_cha  <- economic_cha[-c(53,106,159,212,265,318,371),]


#adding year in each dataframe
economic_cha$year<- c(rep(2010:2017,each=52,length.out=416))


#creation of state variable
state<-c("alabama","alaska", "arizona", "arkansas","california","colorado","connecticut","delaware","district of columbia","florida", "georgia","hawaii" ,"idaho","illinois", "indiana","iowa","kansas","kentucky","louisiana","maine","maryland", "massachusetts","michigan", "minnesota","mississippi","missouri","montana","nebraska","nevada","new hampshire","new jersey","new mexico","new york","north carolina", "north dakota","ohio","oklahoma", 
"oregon","pennsylvania","rhode island","south carolina","south dakota","tennessee","texas","utah","vermont", "virginia","washington", "west virginia","wisconsin" ,"wyoming","puerto rico" )

economic_cha <- cbind(economic_cha,state)

#dropping the state which are not in abortion_dataset
economic_cha<-subset(economic_cha, state!="district of columbia" & state!="puerto rico")

#dropping character space and percentage
economic_cha[,c(1:13)] <- lapply(economic_cha[,c(1:13)], str_replace_all, pattern =  "\\s+", replacement ="")
economic_cha[,c(1:13)] <- lapply(economic_cha[,c(1:13)], str_replace_all, pattern = "%", replacement ="")

#create a dataframe of our variables
economic_cha<-data.frame(economic_cha)

#changing all variable in numeric
economic_cha[,c(1:13)] <- sapply(economic_cha[,c(1:13)], as.numeric)

#renaming all the variable
new_name <- c("eco_empl", "eco_unempl","total_household","eco_inc_h1","eco_inc_h2","eco_inc_h3","eco_inc_h4","eco_inc_h5","eco_inc_h6","eco_inc_h7","eco_inc_h8","eco_inc_h9","eco_inc_h10")
economic_cha<- economic_cha %>% rename_at(1:13,~new_name)

```

***2) Cleaning of the individual characteristics dataset:***
```{r}
#Setting directory
setwd("/Users/coraline/Desktop/Econometrics project")

#Importing dataset individual characteristics
files_i<- c("2010.SelectedIndividualCharacteristics.xlsx","2011.SelectedIndividualCharacteristics.xlsx", "2012.SelectedIndividualCharacteristics.xlsx","2013.SelectedIndividualCharacteristics.xlsx","2014.SelectedIndividualCharacteristics.xlsx", "2015.SelectedIndividualCharacteristics.xlsx", "2016.SelectedIndividualCharacteristics.xlsx", "2017.SelectedIndividualCharacteristics.xlsx")

names<-c("i2010","i2011","i2012","i2013","i2014","i2015","i2016","i2017")

#open all the datasets using a loop
for (i in seq_along(files_i)) {
  assign(names[i], read_excel(path = files_i[i], sheet=2 ))
}


#merging all the datasets
individual_cha<- cbind.data.frame(i2010,i2011,i2012,i2013,i2014,i2015,i2016,i2017)

#transposing data
individual_cha <- data.frame(t(individual_cha), stringsAsFactors=FALSE)
names(individual_cha) <- individual_cha[1,]
individual_cha<- individual_cha[-1,]
individual_cha  <- individual_cha[-c(53,106,159,212,265,318,371),]

#adding year in the dataset
individual_cha$year<-c(rep(2010:2017,each=52,length.out=416))

#selecting variables
individual_cha <-select(individual_cha,"White","Black or African American", "Asian","Hispanic or Latino origin (of any race)", "Never married","Now married, except separated","Divorced or separated","Widowed", "High school graduate (includes equivalency)", "Some college or associate's degree","Bachelor's degree","Graduate or professional degree","Married-couple family","Median Household income (dollars)","year")

#creation of state variable
state<-c("alabama","alaska", "arizona", "arkansas","california","colorado","connecticut","delaware","district of columbia","florida", "georgia","hawaii" ,"idaho","illinois","indiana","iowa","kansas","kentucky","louisiana","maine","maryland", "massachusetts","michigan","minnesota","mississippi","missouri","montana","nebraska","nevada","new hampshire","new jersey","new mexico","new york","north carolina", "north dakota","ohio","oklahoma", "oregon","pennsylvania","rhode island","south carolina","south dakota","tennessee","texas", "utah","vermont", "virginia","washington", "west virginia","wisconsin" ,"wyoming","puerto rico" )

individual_cha <- cbind(individual_cha,state)

#dropping the state which are not in abortion_dataset
individual_cha<-subset(individual_cha, state!="district of columbia" & state!="puerto rico")

#dropping character
individual_cha[,c(1:14)] <- lapply(individual_cha[,c(1:14)], str_replace_all, pattern =  "\\s+", replacement ="")
individual_cha[,c(1:14)] <- lapply(individual_cha[,c(1:14)], str_replace_all, pattern = "%", replacement ="")

#creation of a dataframe
individual_cha<-data.frame(individual_cha)

#changing all variable in numerics
individual_cha[,c(1:14)] <- sapply(individual_cha[,c(1:14)], as.numeric) 

#rename variables
new_name <- c("ind_eth_w","ind_eth_b","ind_eth_a","ind_eth_h","never_married","married", "divorced", "widowed","ind_educ_h","ind_educ_c","ind_educ_b","ind_educ_g","poverty_rate_mariedcouple","median_household_income")

individual_cha<- individual_cha %>% rename_at(1:14,~new_name)


```
***3) Merging the datasets all together***
```{r}
#Merging economic_cha and individual_cha
data <- merge(individual_cha,economic_cha,by=c("year","state"))

#organise rows
data<-data %>% arrange(state)


#label variable
attributes(data$ind_eth_w)$label <- c("Ethnicity: white")
attributes(data$ind_eth_b)$label <- c(" Ehnicity: black or african american")
attributes(data$ind_eth_a)$label <- c(" Ethnicity: asian")
attributes(data$ind_eth_h)$label <- c("Ethnicity: hispanic of latino")
attributes(data$never_married)$label <- c("Marital status: never married ")
attributes(data$married)$label <- c("Marital status:  married ")
attributes(data$divorced)$label <- c("Marital status: divorced or separated ")
attributes(data$widowed)$label <- c("Marital status: widowed ")
attributes(data$ind_educ_h)$label <- c("Level of education: High school graduate ")
attributes(data$ind_educ_c)$label <- c("Level of education: Some college or associate's degree ")
attributes(data$ind_educ_b)$label <- c("Level of education: Bachelor's degree ")
attributes(data$ind_educ_g)$label <- c("Level of education: Graduate or professional degree ")
attributes(data$poverty_rate_mariedcouple)$label<- c("Poverty rate for married couple family")
attributes(data$median_household_income)$label<- c("Median income household (dollars)")
attributes(data$eco_empl)$label <- c("Employment")
attributes(data$eco_unempl)$label <- c("Unemployment")
attributes(data$total_household)$label<- c("Total household")
attributes(data$eco_inc_h1)$label <- c(" Household Income less than $10,000")
attributes(data$eco_inc_h2)$label<- c("Household Income between $10,000 to $14,999")
attributes(data$eco_inc_h3)$label<- c(" Household Income between $15,000 to $24,999")
attributes(data$eco_inc_h4)$label<- c(" Household Income between $25,000 to $34,999")
attributes(data$eco_inc_h5)$label<- c("Household Income between $35,000 to $49,999")
attributes(data$eco_inc_h6)$label<- c("Household Income between $50,000 to $74,999 ")
attributes(data$eco_inc_h7)$label<- c("Household Income between $75,000 to $99,999")
attributes(data$eco_inc_h8)$label<- c("Household income between $100,000 to $149,999")
attributes(data$eco_inc_h9)$label<- c("Household income between $150,000 to $199,999 ")
attributes(data$eco_inc_h10)$label<- c("Household income $200,000 or more")

#compution of the characteristics by women
data <- data%>% mutate_at(vars(ind_eth_w,ind_eth_b,ind_eth_a,ind_eth_h, married, divorced),list(ind_educ_h = ~.*data$ind_educ_h /100))


#Checking for missing values 
skim(data)

```
***3) Creating variables for diff-in-diff design:***

```{r}
#Creation of variables for the diff-in-diff

#creation: treatment variable
summary(data$treatment)
data$treatment <- ifelse((data$state %in% c("alaska", "connecticut", "californie","illinois", "massachusetts","maryland","minnesota","hawaii", "new jersey","new york", "montana","new mexico", "oregon", "vermont","west virginia", "washington")),1,0)

summary(data$treatment)
data$treatment_ <- ifelse((data$state %in% c("alaska", "connecticut", "californie","illinois",  "massachusetts","maryland","minnesota","hawaii", 
"new jersey","new york", "montana","new mexico", "oregon", "vermont","west virginia", "washington")),"Treated","Controlled")

#creation: the post variable
summary(data$post)
data$post = ifelse(data$year >= 2015, 1,0)
data$post_ = ifelse(data$year >= 2015, "After","Before")
summary(data$post)

#keeping variables for the diff-in-diff dataset
data_did <-select(data,"state", "year","married","post","treatment", "divorced")
```


# Part 2: Parallel trends and descriptive statistics

***1) Descriptive statistics***

```{r}
#Summary statistic table 

summarystat_2= data %>% select(never_married, married, divorced, ind_eth_w, ind_eth_b, ind_eth_a, ind_eth_h, ind_educ_g, eco_empl, poverty_rate_mariedcouple, eco_unempl,treatment_, median_household_income ) 

table_2= tbl_summary(summarystat_2,
                     by= treatment_ ,
                     statistic = list(all_continuous() ~ "{mean} ({sd})") , missing = "no", digits = all_continuous() ~ 2,)%>%
  add_n()%>%
  modify_header(label="**Variable**")%>%
  modify_caption("**Summary statistics of the treatment and control group**")%>%
  add_p()
  
table_2
```


```{r}
# Stock of married people 

states_map <- map_data("state")

#Map of the USA 2010 
us_map1 <- subset(data, year=="2010")
names(us_map1)[names(us_map1) == "state"] <- "region"

#Map of the USA 2017
us_map2 <- subset(data, year=="2017")
names(us_map2)[names(us_map2) == "state"] <- "region"


#Calculate percentage change
map_total_usa3 <- merge(us_map1, us_map2, by = "region", suffixes = c("_2010", "_2017"))
map_total_usa3$married_pct_change <- ((map_total_usa3$married_2017 - map_total_usa3$married_2010)/map_total_usa3$married_2010)*100

map_total_usa<- left_join(states_map, map_total_usa3, by= "region")


#Creating the map
map2<- ggplot(map_total_usa, aes(long, lat, group= group))+
  geom_polygon(aes(fill=married_pct_change), color= "white")+ 
  scale_fill_viridis_c()+ theme_bw()+theme(panel.border= element_blank(), panel.grid.major= element_blank(), panel.grid.minor=element_blank())+ labs(
    title = "Percentage Change in Stock of Married People in United States between 2010 and 2017")
map2


```



```{r}
# Stock of divorced people 

states_map <- map_data("state")

#Map of the USA 2010 
us_map3 <- subset(data, year=="2010")
names(us_map3)[names(us_map3) == "state"] <- "region"

#Map of the USA 2017
us_map4 <- subset(data, year=="2017")
names(us_map4)[names(us_map4) == "state"] <- "region"

#Calculate percentage change
map_total_usa5 <- merge(us_map3, us_map4, by = "region", suffixes = c("_2010", "_2017"))
map_total_usa5$divorced_pct_change <- ((map_total_usa5$divorced_2017 - map_total_usa5$divorced_2010)/map_total_usa5$divorced_2010)*100

map_total_usa6<- left_join(states_map, map_total_usa5, by= "region")


#Creating the map
map4<- ggplot(map_total_usa6, aes(long, lat, group= group))+
  geom_polygon(aes(fill= divorced_pct_change), color= "white")+ 
  scale_fill_viridis_c(option = "C")+ theme_bw()+theme(panel.border= element_blank(), panel.grid.major= element_blank(), panel.grid.minor=element_blank())+ labs(
    title = "Percentage Change in Stock of Divorced People in United States between 2010 and 2017")
map4

```

***2) Parallel trends ***
```{r}
#parallel trend marriage 
palette1 = c("#CC79A7", "#009E73")
parallel_trend_1<-ggplot(data, aes(year, married, group=treatment_,color=treatment_)) +
  stat_summary(fun = mean, geom = 'line',aes(linetype =treatment_)) +
  geom_vline(xintercept = 2015) +
  theme_bw()+
  labs( y = 'Stock of married people', 
        x='Year',
        title = 'Figure 3: Parallel trend of the stock of marriage') +
   scale_color_manual(values= palette1)
parallel_trend_1

#parallel trend Divorced
parallel_trend_2<-ggplot(data, aes(year, divorced, group=treatment_,color=treatment_)) +
  stat_summary(fun = mean, geom = 'line',aes(linetype =treatment_)) +
  geom_vline(xintercept = 2015) +
  theme_bw()+
  labs( y = 'Stock of divorced people', 
        x='Year',
        title = 'Figure 4: Parallel trend of the stock of divorced') +
   scale_color_manual(values= palette1)
parallel_trend_2
```
#Part 3: Difference-in-difference regression
***Simple difference-in-difference regression***
```{r}
#regression diff in diff with treatment and post variable : MARRIED
reg_did = lm(married ~ treatment + I(treatment*post) + post, data=data)

#regression diff in diff with treatment and post variable : DIVORCED
reg_did2 = lm(divorced ~ treatment + I(treatment*post) + post, data=data)

```

```{r}
#creation of the table with the results of the regression
 stargazer(reg_did, reg_did2, type = "latex",
          title= "Difference in difference estimates",
          omit=c("year","state"),
          align=TRUE,
          dep.var.labels=c("Stock of marriage", "Stock of divorce"),
          no.space=TRUE)
```


***Difference-in-difference estimation using fixed effects***
```{r}
#regression diff in diff including using fixed effect: MARRIED
reg_did31 = lm(married ~ I(treatment*post) + factor(year), data=data)
reg_did32 = lm(married ~ I(treatment*post) + factor(state), data=data)
reg_did33 = lm(married ~ I(treatment*post) + factor(year) + factor(state), data=data)


reg_did41 = lm(divorced ~ I(treatment*post) + factor(year), data=data)
reg_did42 = lm(divorced ~ I(treatment*post) + factor(state), data=data)
reg_did44 = lm(divorced ~ I(treatment*post) + factor(year) + factor(state), data=data)
```

```{r}
#creation of the table with the results of the regression
 stargazer(reg_did31, reg_did32, reg_did33,type = "latex",
          title= "Difference in difference estimates including fixed effects: Stock of marriage",
          omit=c("year","state"),
          align=TRUE,
          dep.var.labels.include =FALSE,
           column.labels   = c("Years FE", "States FE","Years and states FE"),
          no.space=TRUE)
```
```{r}
#creation of the table with the results of the regression
 stargazer(reg_did41, reg_did42, reg_did44,type = "latex",
          title= "Difference in difference estimates including fixed effects: Stock of divorce",
          omit=c("year","state"),
          align=TRUE,
          dep.var.labels.include =FALSE,
           column.labels   = c("Years FE", "States FE","Years and states FE"),
          no.space=TRUE)
```


***Robustness test***
```{r}
#diff in diff low-income

reg_did8= lm(married ~ treatment + I(treatment*married_ind_educ_h) + post, data=data)
  
reg_did9 =  lm(divorced ~ treatment + I(treatment*divorced_ind_educ_h) + post, data=data)
```

```{r}
#creation of the table with the results of the regression
 stargazer(reg_did8, reg_did9, type = "latex",
          title= "Difference-in-difference estimates focusing on low-income individuals",
          omit=c("year","state"),
          align=TRUE,
          dep.var.labels=c("Stock of marriage", "Stock of divorce"),
          no.space=TRUE)
```
```{r}
#diff in diff adding  controls
reg_did7 = lm(married ~ I(treatment*post) + factor(year) + factor(state)+ ind_eth_b+ ind_eth_a + eco_empl, data=data)

reg_did8= lm(divorced~ I(treatment*post) + factor(year) + factor(state)+ ind_eth_b+ ind_eth_a + eco_empl, data=data)

```

```{r}
#creation of the table with the results of the regression
 stargazer(reg_did7,reg_did8, type = "latex",
          title= "Difference-in-difference estimates including fixed effects and controlls",
          omit=c("year","state"),
          align=TRUE,
          dep.var.labels=c("Stock of marriage", "Stock of divorce"),
          covariate.labels=c("Treatment*Post",
            "Ehnicity: black or african american",
                             "Ethnicity: asian",
                            "Level of employment"),
          no.space=TRUE)
```



